---
title: "Basic features"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
```

## Tidy Data Principles in `robonomistClient`

The `robonomistClient` package follows **tidy data principles** to ensure that the returned datasets are easy to work with and ready for analysis. This approach helps minimize data preparation time, enabling analysts to focus on insights rather than wrangling data. Below are some of the key principles used:

* **Long Format**: Data is returned in long format whenever possible. This means each row represents a single observation, making it ideal for use with popular analysis and visualization tools in R, such as `ggplot2` and the rest of the tidyverse.
* **Value Variable**: In the long format, the measured quantity is captured in a `value` variable that is always numeric. This ensures consistency across datasets and facilitates easy aggregation, summarization, and visualization.
* **Time Variable**: The time dimension is represented by a date variable of class `Date` or date and time of class `POSIXct`. A date variable is positioned consistently in the dataset for ease of use:
  * **First Day of the Period**: For date-based data (e.g., monthly, quarterly), the date represents *the first day of the period*. For example, "March 2024" is represented as `"2024-03-01"`, making it easy to sort and filter chronologically.
* **Variable Arrangement**: Categorical variables, which represent the grouping dimensions of the data (such as region, industry, or demographic group), are positioned **on the left side** of the dataset. Additional attributes related to observations may be positioned on the right side of the `value` variable. This arrangement ensures that categorical identifiers are easily distinguishable from the measurement variables and attributes, improving readability and simplifying analysis.
* **Consistency**: This tidy structure allows the data to seamlessly integrate with typical R workflows and tools for data science, such as `dplyr` and `tidyr`. By using consistent formatting, robonomistClient data can easily be reshaped, filtered, and summarized to suit a wide range of analytical needs.

## Overview of Datasources

The `robonomistClient` package is designed to simplify working with data from a wide variety of datasources. It allows you to easily access data in a standardized format, freeing you from worrying about how to extract and wrangle data from different providers.

To see all available datasources:
```{r}
library(robonomistClient)
datasources()
```

The package keeps the listed datasources automatically up-to-date, ensuring that you always access the most current versions.

```{r, include = FALSE}
data()
```

## Working with Data Tables

All content within `robonomistClient` is organized as data tables, which are typically tibble objects. The `data()` function is a flexible way to explore and retrieve data tables.

### Listing Data Tables

To list all available data tables:

```{r}
data()
```

To browse the tables interactively, you can also open them in the Data Viewer:

```{r, eval = FALSE}
View(data())
```

### Retrieving a Specific Data Table

If you know the specific ID of the table, you can easily retrieve it:

```{r}
df <- data("StatFin/synt/statfin_synt_pxt_12dx.px")
```

To check the version information (vintage) of a data table, use:

```{r}
data_vintage("StatFin/synt/statfin_synt_pxt_12dx.px")
```

## Filtering by Dataset

You can also narrow your search to a specific dataset. For instance, to explore all available tables from the Finnish Tax Administration (Vero):

```{r, eval = FALSE}
View(data("Vero/"))  # Opens Data Viewer with the tables
data("Vero/")        # Lists tables in the consoledata("Vero/")
```

## Searching and Retrieving Data

The data() function is convenient because it combines searching and retrieving data into a single function:

* **Search Mode**: If the argument matches multiple table IDs or titles, data() returns a tibble of matched data tables. For instance:
  ```{r}
  data("väestö")  # Returns multiple matches for 'väestö'
  ```
* **Retrieve Mode**: If the argument matches exactly one table ID, data() will directly return that data table:
  ```{r}
  data("StatFin/synt/statfin_synt_pxt_12dx.px")  # Returns the specific data table
  ```
For more robust usage in production, where consistency is key, it is recommended to use more specific functions:

### `data_search()` and `data_get()`

To clearly separate search and retrieval operations:

* `data_search()`: Use this function to search for matching table IDs without downloading the actual data. This is useful for exploring available datasets:
  ```{r}
  data_search("väestö")
  ```
* `data_get()`: Use this function for robust retrieval of a specific data table by its ID. This ensures either a data table or an error is returned, making it more predictable for production use:
  ```{r}
  d <- data_get("StatFin/synt/statfin_synt_pxt_12dx.px")
  ```

## Handling Time and Labels in Data

### Tidy Time Formatting

For most datasources, `robonomistClient` can automatically format the time dimension for easier analysis. Setting the `tidy_time` argument to `TRUE` will:

* Name the time dimension time.
* Convert it to `Date` class.
* Relocate the time variable to be just before the value column, making it more convenient to work with.

```{r}
# Without tidy time formatting
data_get("StatFin/ntp/statfin_ntp_pxt_132h.px")

# With tidy time formatting
data_get("StatFin/ntp/statfin_ntp_pxt_132h.px", tidy_time = TRUE)
```

The `tidy_time` argument defaults to `TRUE` for all datasources that provide a consistent schema for time variables.

### Working with Labels and Codes

Some datasources provide both labelled and coded versions of data tables. By default, labels are included, but you can control this using the labels argument:

```{r}
# Retrieve data without labels (useful when working with coded data)
data_get("StatFin/ntp/statfin_ntp_pxt_132h.px", labels = FALSE)
```
